---
title: "Check Data M'baiki"
format:
  pdf:
    toc: false
    number-sections: true
    colorlinks: true
execute:
  echo: false      
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
library(data.table)
library(BIOMASS)
library(ggplot2)
library(tidyverse)
```

# Data

This document aims at checking the problem observed in M'Baiki data, showing a breaking point in the trajectories of BA and AGB.

For that, we use the data of M'Baiki on which the data_agregation script has been applied until the chunck *taper-correction*, so that the Diameter have only been corrected with the taper correction.

```{r load-data}
data_cor <- read.csv("data/derived_data/data_Mbaiki_only_taper_cor.csv")
```

I calculate the BA of each tree, and the aggregated value per plot with the uncorrected and corrected diameter.

```{r calc-BA-tree}
# BA of each tree in m²
data_cor <- data_cor |>
  mutate(BA_raw = pi * (Diameter / 200)^2) |>
  mutate(BA_cor = pi * (diameter_cor / 200)^2)
```

```{r calc-BA-plot}
#| message: false
data_aggr <-
  data_cor |>
  group_by(Site, Plot, Year) |>
  summarise(
    BA_raw = sum(BA_raw / PlotArea),
    BA_cor = sum(BA_cor / PlotArea)
  )
```

# Trajectories of BA

Let's look at the trajectories of uncorrected (blue) and corrected (red) BA per plot.

```{r graph_BA}
data_aggr |>
  ggplot(aes(x = Year)) +
  geom_line(aes(y = BA_raw), col = "blue", linewidth = 0.8) +
  geom_line(aes(y = BA_cor),
    col = "red",
    linetype = "dashed", linewidth = 0.8
  ) +
  ylab("BA") +
  theme_minimal() +
  facet_wrap(~Plot)
```

> The taper correction is responsible for the breaking point in the trajectories.

# Number of trees corrected with a taper correction

```{r nb_tapcor}
# count the number of taper corrected trees
n_cor <- data_cor |>
  filter(Diameter != diameter_cor) |>
  group_by(Site, Plot, Year) |>
  count()
# add it to data_aggr
data_aggr <- left_join(data_aggr, n_cor,
  by = c("Site", "Plot", "Year")
)
```

```{r graph_nb_tapcor}
data_aggr$Plot <- as.factor(data_aggr$Plot)
data_aggr |>
  ggplot(aes(x = Year, y = n, colour = Plot, group = Plot)) +
  geom_line() +
  ylab("Number of trees with taper correction") +
  theme_minimal()
# data_aggr |> group_by(Year) |> summarise(Ntot = sum(n))
```

> Taper corrections are made on a high number of trees starting from 2018.

# Explore differences of BA between corrected and uncorrected diameter

I calculate the difference in BA (absolute and relative) due to the taper correction.

```{r calc_corr_effect}
data_cor <- data_cor |>
  mutate(delta_cor = BA_cor - BA_raw) |>
  mutate(delta_cor_rel = delta_cor / BA_raw)
```

```{r}
data_cor |>
  filter(delta_cor != 0) |>
  ggplot(aes(x = HOM, y = delta_cor_rel)) +
  geom_point() +
  ylab("(BA corrected - BA raw) / BA raw") +
  theme_minimal()
```

> The relative correction in BA is correlated to the HOM (expected), but there seems to be a lot of over-corrected values for HOM = 4.5...
